---
title: "AnophelesModel Documentation"
author: "Monica Golumbeanu, Olivier Briët, Clara Champagne, Jeanne Lemant, Barnabas Zogo, Maximilian Gerhards, Marianne Sinka, Nakul Chitnis, Melissa Penny, Emilie Pothin, Tom Smith"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        fig_caption: yes
        toc: true
vignette: >
  %\VignetteIndexEntry{AnophelesModel Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

<style>
p.caption {
  font-size: 0.85em;
  font-style: italic;
  color: grey
}
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
The `AnophelesModel` package can be used to parameterize a model of the
mosquito feeding cycle using data about mosquito bionomics 
(entomological characteristics) and biting 
patterns, as well as human activity and intervention effects. The different 
types of data have been extracted from field studies and are included in 
the package. The model infers the species-specific impact of various
vector control interventions on the vectorial capacity. The package can be used 
to compare the impact of interventions for different mosquito species and to 
generate parameterizations for the entomological and vector control components 
of more complex models of malaria transmission dynamics.

The `AnophelesModel` package is based on a state transition model 
of the feeding dynamics of a mosquito population biting a population of 
hosts defined in [@Chitnis2008]. The model consists of a system of three 
difference equations
describing the dynamics in the numbers of not-infected, infected, and infective
mosquitoes, respectively. The transitions between the individual steps of the 
cycle for a host *i* are modeled through probabilities and 
summarized as follows:

```{r, echo=FALSE, results='asis'}
DiagrammeR::grViz("digraph {
  graph [layout = dot, rankdir = LR]
  
  node [shape = circle, fixedsize=TRUE, fontsize=7]        
  rec1 [label = <Host<BR/>seeking>]
  rec2 [label = <Host<BR/>found>]
  rec3 [label = <Fed>]
  rec4 [label = <Resting>]
  rec5 [label = <Ovipositing>]
  
  # edge definitions with the node IDs
  rec1 -> rec2 [label= <P<SUB>Ai</SUB><BR/> >,fontsize=7]
  rec2 -> rec3 [label= <P<SUB>Bi</SUB><BR/> >,fontsize=7]
  rec3 -> rec4 [label= <P<SUB>Ci</SUB><BR/> >,fontsize=7]
  rec4 -> rec5 [label= <P<SUB>Di</SUB><BR/> >,fontsize=7]
  rec5 -> rec1 [label= <P<SUB>Ei</SUB><BR/> >,fontsize=7]
  }",
  width = 700)
```

These probabilities (represented by the arrows in the diagram above) 
are in turn affected by mosquito ecology, human behavior 
and interventions as described in the following sections of this vignette. 

This documentation provides information about the various use-cases of the
package with examples. First, it gives an overview of the general 
analysis workflow. Next, it presents the contents of the
package database and its data objects. Finally, it provides more detailed 
example use cases of the package functions. These examples include how to 
parameterize the embedded entomological model and evaluate the impact of various
vector control interventions for different mosquito species and geographical 
locations. In addition, examples are shown about how to produce formatted 
inputs for downstream analyses with 
the [OpenMalaria](https://github.com/SwissTPH/openmalaria/wiki) [@Smith2006]
individual-based model of malaria transmission dynamics.

## General workflow
With the main `AnophelesModel()` package function, the user can run in one go 
the main analysis steps for parameterizing the model of the 
mosquito feeding cycle [@Chitnis2008]
and estimate the reduction on the vectorial capacity for various vector control 
interventions. These steps are summarized in the diagram below, alongside the 
corresponding functions:
```{r, echo=FALSE, results='asis'}
DiagrammeR::grViz("digraph {
  graph [layout = dot, rankdir = LR]
  
  node [shape = rectangle]        
  rec1 [label = <Define entomological parameters <BR/><B>def_vector_params()</B>>]
  rec2 [label = <Define mosquito and human activity patterns <BR/><B>def_activity_patterns()</B>>]
  rec3 [label = <Define human host parameters <BR/><B>def_host_params()</B>>]
  rec4 [label = <Parameterize entomological model <BR/><B>build_model_obj()</B>>]
  rec5 [label = <Define interventions effects <BR/><B>def_interventions_effects()</B>>]
  rec6 [label = <Calculate impact on vectorial capacity <BR/><B>calculate_impact()</B>>]
  
  # edge definitions with the node IDs
  rec1 -> rec4 
  rec2 -> rec4 
  rec3 -> rec4
  rec4 -> rec6
  rec5 -> rec6
  }",
  width = 720)
```

The user can directly run the `AnophelesModel()` function as an initial test 
run with a default analysis. By default, the `AnophelesModel()` function 
performs the above steps considering 
the parameters for *An. gambiae* in a Kenyan setting and with default activity 
patterns for mosquitoes and humans. It calibrates the entomological model,
estimates the impact of indoor residual spraying (IRS), long-lasting 
insecticide-treated nets (LLINs) and house screening, then plots the average 
vectorial capacity for a range of intervention coverages:
```{r, dpi=150, fig.cap="Output figure generated with the standard call of the AnophelesModel() function"}
# load the AnophelesModel package
library(AnophelesModel)

# call main function with default values
results_gambiae = AnophelesModel()
```

In the section [Estimating the impact of vector control interventions], 
we describe in detail how the user can access the 
functionalities of the package to customize the 
call to this function or set up their own workflow to run the 
analysis within the settings of interest.

## AnophelesModel database 
The package includes a comprehensive, curated database describing mosquito,
human and intervention characteristics extracted after processing publicly
available field data and organized into multiple data objects.
Precisely, these objects include entomological (or bionomics) parameters
characterizing various *Anopheles* species, their transitions in 
the mosquito feeding cycle,
activity cycles of humans and mosquitoes, as well as parameters 
estimated from experimental hut studies describing
the effects of various interventions.
These parameters are used together within the entomological model included in
the package to estimate the impact of interventions on the vectorial capacity.
All the data objects are directly accessible to the user
after loading the `AnophelesModel` package. 

### Host-specific entomological parameters

The entomological model in the `AnophelesModel` package considers three classes
of hosts (denoted with subscript i): humans protected by interventions, 
non-protected humans,
and animal hosts. Each type of host determines specific transition probabilities
between the consecutive stages of the mosquito feeding cycle
(e.g., interventions such as nets will affect the probability of mosquito
feeding on protected humans). The object `host_ent_param`
in the package database contains, in absence of interventions, for human
and animal hosts, default probabilities that a mosquito
completes the consecutive stages of the feeding cycle in one night,
conditional on having reached each stage:

* `PBi`: Probability that a mosquito bites the host
* `PCi`: Probability that a mosquito finds a resting place after biting
* `PDi`: Probability that a mosquito survives the resting phase
* `PEi`: Probability that a mosquito lays eggs and returns to host-seeking
* `Kvi`: Proportion of susceptible mosquitoes that become infected after biting

The values of the different probabilities in absence of interventions 
are those proposed by [@Chitnis2008] and [@briet2019]. 
The `host_ent_param` object can be directly 
used after loading the `AnophelesModel` package: 
```{r}
# Print the object
host_ent_param
```

### Mosquito bionomics

The object `vec_ent_param` contains bionomic parameterizations for
57 _Anopheles_ mosquito species and 17 complexes (families of species). 
These parameters have been estimated using 
a Bayesian hierarchical model (Lemant et al, in preparation) accounting for 
the phylogeny of 
the _Anopheles_ genus and informed by previously published entomological data 
[@Massey2016], [@briet2019]. Below, we provide an exhaustive list of these 
parameters, their names reflecting the notations used in [@Chitnis2008]:

* `species_name`: name of the mosquito species
* `M`: parous rate (proportion of host seeking mosquitoes that have laid 
eggs at least once)
* `M.sd`: standard deviation of the parous rate 
* `Chi`: human blood index (proportion of mosquito blood meals obtained 
from humans)
* `A0`: sac rate (proportion of mosquitoes who laid eggs the same day)
* `A0.sd`: standard deviation of the sac rate
* `zeta.3`: relative availability of different non-human hosts
* `td`: proportion of a day that a mosquito actively seeks a host
* `tau`: time required for a mosquito that has encountered a host to return to 
host seeking
* `ts`: duration of the extrinsic incubation period (time required for 
sporozoites to develop in mosquitoes)
* `endophily`: proportion of indoor resting mosquitoes
* `endophily.sd`: standard deviation of the proportion of indoor 
resting mosquitoes
* `endophagy`: proportion of indoor feeding mosquitoes
* `endophagy.sd`: standard deviation of the proportion of indoor 
feeding mosquitoes

The `vec_ent_param` object is available upon loading the `AnophelesModel` 
package and can be directly used:
```{r}
# print the first entries of the vec_ent_param object
head(vec_ent_param)
```
For example, to retrieve the parameters for *Anopheles albimanus*:
```{r}
vec_ent_param[vec_ent_param$species_name == "Anopheles albimanus", ]
```

### Human and mosquito activity patterns
The table `activity_patterns` contains mosquito biting patterns as well as 
human sleeping activity information extracted from [@Sherrard-Smith15086] and
[@briet2019]. Each entry has the following attributes:

* `id`: entry ID
* `species`: can be "Homo sapiens" when the entries reflect human activity, 
or names of mosquito species (e.g., "Anopheles gambiae") for mosquito 
biting patterns
* `sampling`: can be one of the following:
    + `IND`: for entries representing the proportion of humans indoors
    + `BED`: for entries representing the proportion of humans in bed
    + `HBI`: for entries representing the proportion of indoors human 
    biting
    + `HBO`: for entries representing the proportion of outdoors human 
    biting
    + `ABO`: for entries representing the proportion of outdoors 
    animal biting
    + `HB`: for entries corresponding to the proportion of human biting
* `country`: country where the measurements were taken
* `site`: name of the geographical site where the measurements were taken
* `hour`: hour of the day for which the sampling is reported, ranging from 4pm 
until 8am, format: hh:mm_hh:mm (e.g., 16:00_17:00)
* `value`: measurement value

The object can be directly examined and its components
retrieved for various purposes. For example, to retrieve all the human 
activity patterns collected from Kenya:
```{r}
select_idx = activity_patterns$species == "Homo sapiens" &
                                  activity_patterns$country == "Kenya"
Kenya_human_rhythms = activity_patterns[select_idx,]
head(Kenya_human_rhythms)
```
A pattern of interest can then be selected and displayed, for example:
```{r, dpi=150, fig.cap="Example of human activity patterns in Rachuonyo, Kenya. The proportion of humans indoors (labeled IND) out of the total human population is displayed."}
library(ggplot2)
Rachuonyo_human_rhythms = Kenya_human_rhythms[which(Kenya_human_rhythms$id == 2),]
labels = substr(unique(Rachuonyo_human_rhythms$hour), 1, 5)
Rachuonyo_human_rhythms$hour = factor(labels, levels = labels )
ggplot(Rachuonyo_human_rhythms, aes(x=hour, y=value, group=site)) +
    theme_light() + theme_linedraw() + theme_bw(base_size = 5) +
    theme(legend.position = "bottom") +
    geom_line(size=0.5) + 
    scale_x_discrete(name = "Time of the day (hh.mm)")+ 
    guides(color = guide_legend(nrow = 2,byrow = TRUE)) +
    labs(title = "Proportion of humans indoors in Rachuonyo, Kenya", fill="",
         x="Time", y="Activity")
```

Another example extracting all the available biting behavior for 
*Anopheles gambiae* in Kenya:
```{r}
select_idx = activity_patterns$species == "Anopheles gambiae" &
                                  activity_patterns$country == "Kenya"
Kenya_gambiae_biting = activity_patterns[select_idx,]
head(Kenya_gambiae_biting)
```

### Intervention parameters 
The `AnophelesModel` package includes parameterisations for estimating the 
effects of various vector control interventions. These effects include
reducing host availability, killing mosquitoes over time since spraying,
insecticidal and deterrent effects as functions of insecticide content. 
The interventions modeled in the package consist of different types 
of long lasting insecticide treated nets (LLINs), 
indoor residual spraying (IRS), and house screening. All the parameters for 
calculating these effects are provided within the `interventions_param` 
object which consists of three components, namely:

* `interventions_param$interventions_summary`: an overview and general 
information for all available intervention parameterisations; this includes 
information about the active agent (insecticide) used, the 
mosquito species for which the effects were assessed in the experimental trials, 
the duration of the intervention and details about the method and references
* `interventions_param$LLINs_params`: an object including parameterisations 
for various types of LLINs
* `interventions_param$IRS_params`: a table containing estimated insecticide 
decay for various types of IRS

The effects of LLINs are parameterized in terms of effectiveness in 
reducing the availability of humans, and both pre- and post-prandial killing 
of mosquitoes respectively. The effects are based on estimates from 
experimental hut studies from [@randriamaherijaona2015]. The decay of these 
effects over time, in terms of attrition, use, physical and chemical integrity 
is parameterized through a series of logistic models described in [@briet2013], 
[@briet2019], and [@briet2020] using the data of President Malaria Initiative 
(PMI) net durability studies (7 countries, 8 net types, 23 combinations in all), 
and also from [@morgan2015]. The object `interventions_param$LLINs_params` 
contains all this information and is composed of three tables: 

* `interventions_param$LLINs_params$model_coeff`: values for coefficients of 
regression models estimating LLINs effects; these models have been defined in 
[@briet2013], [@briet2019] and [@briet2020]
* `interventions_param$LLINs_params$insecticide_c`: insecticide content for 
various types of nets summarized in [@briet2020]
* `interventions_param$LLINs_params$durability_estim`: information about the 
durability of LLINs (holed area, survival) across time in different 
geographical locations also summarized in [@briet2020].

Similarly, for IRS, a table with time series of effects in reducing human 
availability and killing of mosquitoes as well as parameters describing 
decay of effects over time is provided in `interventions_param$IRS_params`. 

The parameterization for House screening is based 
on the estimates from [@kirby2009] and described in [@briet2019], assuming a 
house entry reduction of 59%.

The data objects for interventions described above are directly accessible upon 
loading the `AnophelesModel` package. For example, for obtaining an overview
of all LLINs interventions provided with the `AnophelesModel` package:

```{r}
print(interventions_param$interventions_summary[which(interventions_param$interventions_summary$Intervention == "LLINs"),])
```

### Functions querying the package database

The package additionally contains helper functions for listing an overview 
of the different data objects to help the user populate the input arguments of 
the key functions in the general workflow (e.g., for the `AnophelesModel()` 
function). These are:

* `list_all_species()`: prints all the mosquito species for which bionomic 
parameters are provided in the package database
* `list_activity()`: prints all the available activity patterns 
* `list_interventions()`: prints the available vector control interventions 
* `list_intervention_models()`: prints the available intervention 
parameterisation IDs
* `get_net_types()`: prints the available net types and the countries where 
the studies have been conducted 

Furthermore, the package allows extracting more specific information about the 
durability of different types of LLINs such as holed area, survival and decay. 
These parameteres were estimated in the PMI net durability studies.

For example, to retrieve the decay characteristics 
for the DuraNet LLIN in Kenya:
```{r}
# load the necesary packages
library(reshape)

DuraNet = get_net_decay(net_type = "DuraNet", country = "Kenya", insecticide_type = "DuraNet", n_ips = 100, duration = 3)
```
Then, to plot these characteristics :
```{r, fig.show='hold'}
library(ggpubr)
# Plot the survival
ggplot(DuraNet, aes(x=time, y=survival)) + geom_point() +
    theme_light() + theme_linedraw() + theme_bw() + ggtitle("Survival of the net")

# Plot the log of the holed area
ggplot(DuraNet, aes(x=time, y=logHoles)) + geom_point() +
    theme_light() + theme_linedraw() + theme_bw() + ggtitle("Log holed area")

# Plot the insecticide content
ggplot(DuraNet, aes(x=time, y=insecticideContent)) + geom_point() +
    theme_light() + theme_linedraw() + theme_bw() + ggtitle("Insecticide decay")
```

## Estimating the impact of vector control interventions

The mosquito bionomics, feeding cycle transitions, 
activity patterns and intervention effects accessible from 
the `AnophelesModel` database can be used to parameterise the model of 
the mosquito feeding cycle
and estimate the impact of interventions on the vectorial capacity.

### Step 1: Defining entomological parameters

To define the mosquito specific entomological parameters in absence of 
interventions, the user can either 
select existing parameterisations from the `AnophelesModel` database, or 
provide a set of custom values. For example, to choose an existing 
parameterisation for *Anopheles nili*:

```{r}
nili_ent_params = def_vector_params(mosquito_species = "Anopheles nili")
head(nili_ent_params)
```

To use a different parameterisation, first all the entomological parameter values need 
to be provided in a data frame with the same column names 
as the `vec_ent_param` object:
```{r}
# Use own parameterisation
species_name = "Anopheles example"
M = 0.623
M.sd = 0
Chi = 0.939
A0 = 0.313
A0.sd = 0
zeta.3 = 1
td = 0.33
tau = 3
ts = 10
to = 5
endophily = 1
endophily.sd = 0
endophagy = 1
endophagy.sd = 0
custom_ent_params_table = as.data.frame(cbind.data.frame(species_name, M, M.sd, Chi,
                                A0, A0.sd, zeta.3, td,
                                tau, ts, to, endophily, endophily.sd, endophagy, endophagy.sd))
```
Then, the data frame is provided to the function defining the vector parameters:
```{r}
custom_ent_params = def_vector_params(mosquito_species = species_name, vector_table = custom_ent_params_table)
head(custom_ent_params)
```

### Step 2: Defining host-specific entomological parameters
Objects for host-specific entomological parameters (e.g., biting of human 
versus animal hosts, for details or other parameters see section 
[Host-specific entomological parameters]) in absence of interventions 
can be defined in a similar way as 
for the mosquito bionomics. To choose the default parameterisation 
from [@Chitnis2008]:
```{r}
default_host_params = def_host_params()
print(default_host_params)
```

For using a different, custom parameterisation, first we need to build a 
data frame with the same structure as `host_ent_params`:
```{r}
species_name = "Anopheles example" 
host = c("human", "animal") 
PBi = c(0.67, 0.44)
PCi = c(0.33, 0.434)
PDi = c(0.647, 0.344)
PEi = c(0.672, 0.214)
Kvi = c(0.52, 0.314)
custom_params_tab = cbind.data.frame(species_name, host, PBi, PCi, PDi, PEi, Kvi)
```
Afterwards, we can provide this data frame to the relevant function for defining 
the host parameters:
```{r}
custom_host_params = def_host_params(mosquito_species = species_name, host_table = custom_params_tab)
print(custom_host_params)
```

### Step 3: Defining mosquito and human activity patterns
To facilitate definition of human and mosquito activity patterns, the 
`AnophelesModel` package comes with inbuilt default values for 
_Anopheles gambiae_ in Kenya and _Anopheles albimanus_ in Haiti. 
The user can choose other patterns from the package database (344 entries) by 
using the entry ID of each time series (`id` column of object 
`activity_patterns`). Furthermore, customized patterns can be defined by 
directly providing the time series values. Here are some examples for 
these different use cases:

To use default activity patterns from Kenya (human activity measured in the 
Rachuonyo and Rarieda districts):
```{r}
activity_p = def_activity_patterns(activity = "default_Anopheles_gambiae")
print(activity_p)
```

To select activity patterns from the database based on the entry IDs for 
*Anopheles arabiensis* in Kenya, Ahero disctrict:
```{r}
activity_list = NULL
activity_list$HBI = 34
activity_list$HBO = 165
activity_list$humans_indoors = 4
activity_list$humans_in_bed = 24
activity_p2 = def_activity_patterns(activity_list)
print(activity_p2)
```

To use custom activity patterns by specifying the different time series, for
example for *An. albimanus* in Haiti available from [@briet2019]:
```{r}
# Briet et al 2019, Table S5. Biting rhythm of An. albimanus in Haiti, Dame Marie
HBI = c(0.22, 0.21, 0.22, 0.10, 0.13, 0.17, 0.08, 0.12, 0.03, 0.12, 0.17, 0.18, 0.25)
HBO = c(0.25, 0.35, 0.37, 0.33, 0.36, 0.32, 0.13, 0.14, 0.09, 0.15, 0.36, 0.23, 0.25)
# Human activity in Haiti from Briet et al 2019, Knutson et al. 2014
humans_in_bed = c(5.628637, 14.084496, 28.558507, 47.761203, 67.502964, 83.202297, 92.736838, 96.734455, 96.637604, 92.983136, 85.264515, 73.021653, 57.079495)/100 
humans_indoors = c(5.628637, 14.084496, 28.558507, 47.761203, 67.502964, 83.202297, 92.736838, 96.734455, 96.637604, 92.983136, 85.264515, 73.021653, 57.079495)/100 
custom_params = as.data.frame(cbind(HBI, HBO, humans_indoors,humans_in_bed ))
custom_activity_obj = def_activity_patterns(custom_params)
print(custom_activity_obj)
```

### Step 4: Initializing the entomological model
Once the vector, host and activity parameters have been defined, the 
entomological model can be initialized and calibrated accordingly.
In this step, the death rate and the availability to mosquitoes 
for each host type are estimated. For a description of how these two 
parameters are estimated, see [@Chitnis2008].

First, the sizes of the host and mosquito populations need to be defined:
```{r}
# Define size of the host and mosquito population 
host_pop = 2000
vec_pop = 10000
```

Then, to initialize the entomological model and build the model object:
```{r}
model_params = build_model_obj(custom_ent_params, default_host_params, custom_activity_obj, host_pop)
# Print the calculated mosquito death rate
print(model_params$host_params$muvA)
# Print the estimated availability to mosquitoes for each host type
print(model_params$host_params$alphai)
```

### Step 5: Defining interventions effects
In the `AnophelesModel` package, the interventions and their characteristics 
are specified through a list of intervention objects. A detailed description of
how the intervention objects are specified can be obtained by using the command 
`?def_interventions_effects`. To facilitate definition of these objects, 
the package contains a list of intervention object examples 
for each intervention type included in the `intervention_obj_examples` list. 

The `intervention_obj_examples` list contains three examples of interventions :
LLIN, IRS and House screening, respectively: 
```{r}
print(intervention_obj_examples)
```
Except for the house screening where only the `id` and `description` attributes 
are required, to use the IRS and LLIN interventions included in the package, 
the user needs to 
provide the parameterisation ID (e.g., LLINs01, IRS01). A list of all package
intervention parameterisations and their IDs can be obtained with the 
`list_intervention_models()` command.
The examples from `intervention_obj_examples` can be adapted to the user's 
needs to pick other available parameterisations and intervention 
characteristics available within the package. For example, to select a 
different IRS parameterisation, we can modify the `parameterisation` attribute 
of the IRS example:
```{r}
new_IRS = intervention_obj_examples$IRS_example
new_IRS$description = "Permethrin IRS"
new_IRS$parameterisation = "IRS16"
```
then we can concatenate the new intervention object to the list with 
intervention examples:
```{r}
new_intervention_list = c(intervention_obj_examples, list(new_IRS = new_IRS))
```

Once the intervention parameterisations have been defined, 
the user needs to specify the intervention deployment coverages and a number of 
time points. The duration of each intervention will be evenly split across the 
time points and the intervention effects will be calculated at each time point.
```{r}
list_interv = new_intervention_list
coverages = c(seq(0, 1, by = 0.1))
n_ip = 100
# Calculate the intervention effects:
intervention_vec = def_interventions_effects(list_interv, model_params, n_ip)
```

The effects of the interventions are adjusted according to indoor and outdoor 
human exposure to mosquitoes. The user can calculate the human exposure 
parameters using the following function:
``` {r}
in_out_exp = get_in_out_exp(activity_cycles = custom_activity_obj, vec_p = custom_ent_params)
print(in_out_exp)
```

### Step 6: Calculating intervention impact on vectorial capacity
Once all the intervention effects have been defined, their impact on the 
vectorial capacity can be estimated (using equation 20 from [@Chitnis2008]) :
```{r, dpi=150, fig.cap="Output figure with the impact on the vectorial capacity for different species"}
impacts = calculate_impact(intervention_vec, coverages, model_params,
vec_pop, n_ip)
```
The resulting vectorial capacity is calculated and aggregated across the 
previously-specified time points which are uniformly-distributed across the 
duration of an 
intervention. To display the average vectorial capacity for various coverages:
```{r, dpi=150, fig.cap="Impact of interventions on vectorial capacity."}
p = plot_impact_species(impacts, "VC_red")
plot(p)
```

The variability in the entomological parameters for each mosquito species can 
be accounted for when calculating the impact of interventions on the vectorial 
capacity. To do so, the standard deviation values of the entomological 
parameters (namely parameters `M.sd`, `A0.sd`, `endophily.sd` and `endophagy.sd` 
in the `vec_ent_param` object) 
are used to create a confidence interval defined as the mean value 
+/- two standard deviations. Using latin hypercube sampling [@stein1987], 
a set of samples are uniformly drawn from the 
multi-dimensional parameter space defined by the confidence intervals. Then, 
the average vectorial capacity is computed for each of these samples and the 
resulting confidence interval for the vectorial capacity is represented by the 
mean value +/- two standard deviations. This sampling-based approach is 
included in the `calculate_impact_var()` function to 
estimate the confidence interval for the reported mean vectorial capacity. 
In the example below, in the interest of execution time, only 10 samples have 
been used to estimate the confidence intervals of the vectorial capacity:
```{r, dpi=150, fig.cap="Impact of interventions on vectorial capacity including the variability in the entomological parameters specific to the mosquito species."}
impact_gambiae = calculate_impact_var(mosquito_species = "Anopheles gambiae",
                                        activity_patterns = "default_Anopheles_gambiae",
                                        interventions = intervention_obj_examples,
                                        n_sample_points = 10,
                                        plot_result = FALSE)
plot_impact_var("Anopheles gambiae", impact_gambiae)
```

## Generating vector parameterisations for OpenMalaria
The package `AnophelesModel` can be used to create XML snippets containing
parameterisations for the entomology and generic vector control interventions 
(GVI) components of OpenMalaria experiments. 
These snippets can then be incorporated in the base 
XML files describing the setup for OpenMalaria individual-based simulations.
There are two main categories of XML snippets that can be generated with the 
package:

* `<mosq>` and `<nonHumanHosts>` snippets containing entomological parameters
* `<GVI>` snippets with parameters describing the effects of interventions

For example, to generate the `<mosq>` and `<nonHumanHosts>` snippets 
for *Anopheles nili*:
```{r}
entomology_xml = get_OM_ento_snippet(nili_ent_params, default_host_params)
print(entomology_xml)
```

To generate the `<GVI>` snippets with deterrency, pre- and post-prandial effects
of interventions, a set of 7 decay functions are fitted to the time series with 
intervention effects calculated previously with `def_interventions_effects()` 
and the best fit is chosen. The package also produces visualizations of the 
different decay fits. For a detailed description of the fitted decay functions, 
check the [description in the OpenMalaria wiki](https://github.com/SwissTPH/openmalaria/wiki/ModelDecayFunctions).
```{r, dpi=150}

GVI_snippets = get_OM_GVI_snippet("Anopheles example", impacts$interventions_vec$LLINs_example,
                    100, plot_f = TRUE)
```

## References


